
## View of Data

```{r}
kable(head(power_plant_data, format = "html"))
```

```{r}
# Glimpse and Summary
glimpse(power_plant_data)
summary(power_plant_data)
```

## Variable Descriptions for the Global Power Plant Database

1. **`country`**  
   *Description*: The country's code where the power plant is located.  
   *Type*: Categorical (e.g., "AFG", "IND", "GBR")

2. **`country_long`**  
   *Description*: The country's name where the power plant is located.  
   *Type*: Categorical (e.g., "Afghanistan", "India", "	United Kingdom")

3. **`name`**  
   *Description*: The name of the power plant. This typically refers to the formal name assigned to the plant.  
   *Type*: Categorical (e.g., "Vindhyachal Thermal Power Station")

4. **`gppd_idnr`**  
   *Description*: A unique identifier for the power plant in the dataset, useful for cross-referencing.  
   *Type*: Categorical (alphanumeric code)
   
The result of `length(power_plant_data$gppd_idnr) == nrow(power_plant_data)` being `TRUE` confirms that each power plant in the dataset has a unique `gppd_idnr`, making it a reliable identifier for cross-referencing.

5. **`capacity_mw`**  
   *Description*: The capacity of the power plant in megawatts (MW). This indicates the maximum output the plant can generate.  
   *Type*: Numeric (e.g., 1000 MW)

6. **`latitude`**  
   *Description*: The geographical latitude of the power plant’s location.  
   *Type*: Numeric (e.g., 28.7041)

7. **`longitude`**  
   *Description*: The geographical longitude of the power plant’s location.  
   *Type*: Numeric (e.g., 77.1025)

8. **`primary_fuel`**  
   *Description*: The primary fuel used by the power plant to generate electricity. Examples include coal, natural gas, hydro, solar, etc.  
   *Type*: Categorical (e.g., "Coal", "Hydro", "Solar")

9. **`other_fuel1`**  
   *Description*: A secondary fuel used by the power plant, if applicable.  
   *Type*: Categorical (e.g., "Natural Gas")

10. **`other_fuel2`**  
    *Description*: A tertiary fuel used by the power plant, if applicable.  
    *Type*: Categorical (e.g., "Oil")

11. **`other_fuel3`**  
    *Description*: A fourth fuel type, if applicable.  
    *Type*: Categorical (e.g., "Biomass")

12. **`commissioning_year`**  
    *Description*: The year the power plant was commissioned and began operations.  
    *Type*: Numeric (e.g., 1995)

13. **`owner`**  
    *Description*: The owner(s) of the power plant, which could be a private company, a government entity, or a consortium of both.  
    *Type*: Categorical (e.g., "NTPC", "State Government of Maharashtra")

14. **`source`**  
    *Description*: The source from which the data about the power plant was obtained.  
    *Type*: Categorical (e.g., "World Bank", "IEA")

15. **`url`**  
    *Description*: A URL link to more detailed information about the power plant (e.g., official site, reports).  
    *Type*: URL (e.g., "https://www.ntpc.co.in/")

16. **`geolocation_source`**  
    *Description*: The source of the geolocation data, providing information about the geographic coordinates of the power plant.  
    *Type*: Categorical (e.g., "Google Earth", "OpenStreetMap")

17. **`wepp_id`**  
    *Description*: A unique identifier for the power plant in the WEPP (World Energy Project Database).  
    *Type*: Categorical (alphanumeric code)

18. **`year_of_capacity_data`**  
    *Description*: The year when the capacity data was last updated or reported.  
    *Type*: Numeric (e.g., 2019)

19. **`generation_gwh_2013`**  
    *Description*: The electricity generation in gigawatt-hours (GWh) for the year 2013.  
    *Type*: Numeric (e.g., 1200 GWh)

20. **`generation_gwh_2014`**  
    *Description*: The electricity generation in gigawatt-hours (GWh) for the year 2014.  
    *Type*: Numeric (e.g., 1250 GWh)

21. **`generation_gwh_2015`**  
    *Description*: The electricity generation in gigawatt-hours (GWh) for the year 2015.  
    *Type*: Numeric (e.g., 1300 GWh)

22. **`generation_gwh_2016`**  
    *Description*: The electricity generation in gigawatt-hours (GWh) for the year 2016.  
    *Type*: Numeric (e.g., 1350 GWh)

23. **`generation_gwh_2017`**  
    *Description*: The electricity generation in gigawatt-hours (GWh) for the year 2017.  
    *Type*: Numeric (e.g., 1400 GWh)

24. **`generation_gwh_2018`**  
    *Description*: The electricity generation in gigawatt-hours (GWh) for the year 2018.  
    *Type*: Numeric (e.g., 1450 GWh)

25. **`generation_gwh_2019`**  
    *Description*: The electricity generation in gigawatt-hours (GWh) for the year 2019.  
    *Type*: Numeric (e.g., 1500 GWh)

26. **`generation_data_source`**  
    *Description*: The source of the generation data, indicating where the reported generation figures come from.  
    *Type*: Categorical (e.g., "Plant Records", "Estimated")

27. **`estimated_generation_gwh_2013`**  
    *Description*: The estimated electricity generation in gigawatt-hours (GWh) for the year 2013, if actual data is unavailable.  
    *Type*: Numeric (e.g., 1150 GWh)

28. **`estimated_generation_gwh_2014`**  
    *Description*: The estimated electricity generation in gigawatt-hours (GWh) for the year 2014, if actual data is unavailable.  
    *Type*: Numeric (e.g., 1200 GWh)

29. **`estimated_generation_gwh_2015`**  
    *Description*: The estimated electricity generation in gigawatt-hours (GWh) for the year 2015, if actual data is unavailable.  
    *Type*: Numeric (e.g., 1250 GWh)

30. **`estimated_generation_gwh_2016`**  
    *Description*: The estimated electricity generation in gigawatt-hours (GWh) for the year 2016, if actual data is unavailable.  
    *Type*: Numeric (e.g., 1300 GWh)

31. **`estimated_generation_gwh_2017`**  
    *Description*: The estimated electricity generation in gigawatt-hours (GWh) for the year 2017, if actual data is unavailable.  
    *Type*: Numeric (e.g., 1350 GWh)

32. **`estimated_generation_note_2013`**  
    *Description*: A note explaining any discrepancies or assumptions made in the estimated generation for 2013.  
    *Type*: Categorical (e.g., "Data not available")

33. **`estimated_generation_note_2014`**  
    *Description*: A note explaining any discrepancies or assumptions made in the estimated generation for 2014.  
    *Type*: Categorical (e.g., "Data estimated based on plant capacity")

34. **`estimated_generation_note_2015`**  
    *Description*: A note explaining any discrepancies or assumptions made in the estimated generation for 2015.  
    *Type*: Categorical (e.g., "Data estimated")

35. **`estimated_generation_note_2016`**  
    *Description*: A note explaining any discrepancies or assumptions made in the estimated generation for 2016.  
    *Type*: Categorical (e.g., "Actual data unavailable")

36. **`estimated_generation_note_2017`**  
    *Description*: A note explaining any discrepancies or assumptions made in the estimated generation for 2017.  
    *Type*: Categorical (e.g., "Based on historical averages")

This dataset provides a comprehensive overview of the world's power plants, making it an invaluable resource for energy analysis, decision-making, and reporting purposes.

## Missing data rate for different columns

```{r}
# Missing data rate for different columns
missing_data_rate <- power_plant_data %>%
  summarise_all(~ mean(is.na(.)) * 100) %>% # Calculate percentage of missing values
  pivot_longer(everything(), names_to = "column", values_to = "missing_rate") %>%
  arrange(desc(missing_rate))

# View missing data rates
print(missing_data_rate)
```

The column `other_fuel3` has a missing rate of 100%, indicating that no records in the dataset contain information about a fourth fuel type. This could indicate that none of the power plants in the dataset utilizes a fourth fuel type. This suggests that the column is effectively unused and could potentially be excluded from the analysis.


```{r}
# Missing data rate
p <- ggplot(missing_data_rate, aes(x = reorder(column, -missing_rate), y = missing_rate, fill = missing_rate)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis() +
  coord_flip() +
  labs(
    title = "Missing Data Rate by Column",
    x = "Column",
    y = "Missing Data Rate (%)"
  )

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```

## Number of Power Plants by Country

```{r}
# Distribution of the number of power plants by country
plants_by_country <- power_plant_data %>%
  group_by(country_long) %>%
  summarise(num_power_plants = n()) %>%
  arrange(desc(num_power_plants))

# View the distribution
print(plants_by_country)

# 1. Number of power plants by country
p <- ggplot(head(plants_by_country, 10), aes(x = reorder(country_long, -num_power_plants), y = num_power_plants)) +
  geom_bar(stat = "identity", fill = viridis(1)) + # fill = "steelblue"
  coord_flip() +
  labs(title = "Top 10 Country by Number of Power Plants", x = "Country", y = "Number of Power Plants")

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```

## Number of Power plants by primary energy source

```{r}
# Distribution of power plants by primary energy source (primary_fuel)
plants_by_fuel <- power_plant_data %>%
  group_by(primary_fuel) %>%
  summarise(num_power_plants = n()) %>%
  arrange(desc(num_power_plants))

# View the distribution
print(plants_by_fuel)

# Power plants by primary energy source
p <- ggplot(plants_by_fuel, aes(x = reorder(primary_fuel, -num_power_plants), y = num_power_plants, fill = primary_fuel)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_viridis(discrete = TRUE) +
  labs(title = "Number of Power Plants by Primary Energy Source", x = "Primary Energy Source", y = "Number of Power Plants")

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```

## Number of power plants by commissioning year

```{r}
# Number of power plants by commissioning year
plants_by_year <- power_plant_data %>%
  filter(!is.na(commissioning_year)) %>%
  group_by(commissioning_year) %>%
  summarise(num_power_plants = n()) %>%
  arrange(commissioning_year)

# View the yearly distribution
print(plants_by_year)

# Visualization of Number of power plants by commissioning year
p <- ggplot(plants_by_year, aes(x = commissioning_year, y = num_power_plants)) +
  geom_line(color = "blue") +
  geom_point(color = "red") +
  labs(title = "Number of Power Plants by Commissioning Year", x = "Year", y = "Number of Power Plants")

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```



## Power Plant Location Map

```{r}
# 2. 
# Ensure dataset has valid lat/lon values
valid_locations <- power_plant_data %>%
  filter(!is.na(latitude) & !is.na(longitude))

leaflet(data = valid_locations) %>%
  addTiles() %>%
  addCircleMarkers(
    ~longitude, ~latitude,
    popup = ~paste(name, "<br>", "Capacity:", capacity_mw, "MW", "<br>", "Fuel Type:", primary_fuel),
    radius = ~log(capacity_mw + 1), # Scale radius by log of capacity for better visualization
    color = "blue",
    fillOpacity = 0.7
  ) %>%
  addLegend(
    position = "bottomright",
    title = "Power Plants",
    colors = "blue",
    labels = "Power Plant Locations"
  )
```

## Plant Density

```{r}
# Ensure bins for latitude and longitude are numeric for mapping
region_density <- power_plant_data %>%
  filter(!is.na(latitude) & !is.na(longitude)) %>%
  mutate(
    lat_bin_start = floor(latitude / 5) * 5,
    lon_bin_start = floor(longitude / 5) * 5,
    lat_bin_end = lat_bin_start + 5,
    lon_bin_end = lon_bin_start + 5
  ) %>%
  group_by(lat_bin_start, lon_bin_start, lat_bin_end, lon_bin_end) %>%
  summarise(num_plants = n(), .groups = 'drop') %>%
  arrange(desc(num_plants))

# Plot density regions on a Leaflet map
leaflet(data = region_density) %>%
  addTiles() %>%
  addRectangles(
    lng1 = ~lon_bin_start,
    lat1 = ~lat_bin_start,
    lng2 = ~lon_bin_end,
    lat2 = ~lat_bin_end,
    fillColor = ~colorNumeric("viridis", num_plants)(num_plants),
    fillOpacity = 0.7,
    weight = 0
  ) %>%
  addLegend(
    "bottomright",
    pal = colorNumeric("viridis", region_density$num_plants),
    values = region_density$num_plants,
    title = "Plant Density",
    opacity = 1
  )

```

## Geographic Areas Most Equipped with Power Plants

```{r}
# Summarize the number of power plants by geographic region
geographic_density <- power_plant_data %>%
  filter(!is.na(latitude) & !is.na(longitude)) %>%
  mutate(
    lat_bin = floor(latitude / 5) * 5,
    lon_bin = floor(longitude / 5) * 5
  ) %>%
  group_by(lat_bin, lon_bin) %>%
  summarise(num_plants = n(), total_capacity = sum(capacity_mw, na.rm = TRUE), .groups = 'drop')

# Visualization: Map of Power Plant Density
leaflet(geographic_density) %>%
  addTiles() %>%
  addCircles(
    lng = ~lon_bin + 2.5, lat = ~lat_bin + 2.5,
    weight = 1,
    radius = ~sqrt(num_plants) * 10000, # Scale by sqrt of the number of plants
    color = "blue",
    fillColor = "blue",
    fillOpacity = 0.5,
    popup = ~paste0("Plants: ", num_plants, "<br>Capacity (MW): ", total_capacity)
  ) %>%
  addLegend(
    position = "bottomright",
    colors = "blue",
    labels = "Density of Power Plants",
    title = "Power Plant Density"
  )
```

