## Evolution of Installed Capacity by Country Over Time

Understanding the historical growth of installed capacity helps evaluate the development of energy infrastructure in different countries. The following line chart shows the evolution of installed capacity (in MW) for all countries in the dataset.

```{r}
# Evolution of Installed Capacity by Country Over Time
capacity_evolution <- power_plant_data %>%
  filter(!is.na(commissioning_year) & !is.na(capacity_mw)) %>%
  group_by(country, commissioning_year) %>%
  summarise(total_capacity = sum(capacity_mw, na.rm = TRUE)) %>%
  arrange(commissioning_year)

print(capacity_evolution)

p <- ggplot(capacity_evolution, aes(x = commissioning_year, y = total_capacity, color = country)) +
  geom_line() +
  scale_color_viridis(discrete = TRUE, option = "D") +
  labs(
    title = "Evolution of Installed Capacity by Country",
    x = "Commissioning Year",
    y = "Total Installed Capacity (MW)",
    color = "Country"
  ) +
  theme_minimal()

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```
## Energy Production Trends by Primary Fuel Type

```{r}
# Aggregate production data by fuel type and year
production_trends <- power_plant_data %>%
  select(primary_fuel, matches("^generation_gwh_\\d+$")) %>%
  pivot_longer(cols = matches("^generation_gwh_\\d+$"), 
               names_to = "year", 
               values_to = "production") %>%
  mutate(
    year = as.numeric(gsub("generation_gwh_", "", year))
  ) %>%
  group_by(primary_fuel, year) %>%
  summarise(
    total_production = sum(production, na.rm = TRUE),
    .groups = 'drop'
  )

# Visualization: Production Trends by Fuel Type
p <- ggplot(production_trends, aes(x = year, y = total_production, color = primary_fuel)) +
  geom_line(size = 1) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    title = "Energy Production Trends by Fuel Type",
    x = "Year",
    y = "Total Production (GWh)",
    color = "Fuel Type"
  ) +
  theme_minimal()

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```

## Analyze progress in installed capacity and electricity production

To analyze progress in installed capacity and electricity production, we can calculate the percentage growth or absolute differences over a range of years for each country. Then, we can identify countries with the most significant improvements.

**Approach**:

1. **Installed Capacity Progress**:
   - Calculate the difference in total installed capacity (`capacity_mw`) between earlier and later years.
   - Compute percentage growth per country.

2. **Electricity Production Progress**:
   - Sum the annual electricity generation for each country.
   - Compare total production across available years.

3. **Top Performing Countries**:
   - Rank countries by absolute or percentage growth in installed capacity and electricity production.

4. **Visualization**:
   - Use bar charts or line plots to showcase progress for the top-performing countries.

**Code Implementation**:

```{r}
# Calculate progress in installed capacity
capacity_progress <- power_plant_data %>%
  group_by(country_long) %>%
  summarise(
    initial_capacity = sum(capacity_mw, na.rm = TRUE),
    recent_capacity = sum(if_else(commissioning_year >= 2015, capacity_mw, 0), na.rm = TRUE),
    capacity_growth = recent_capacity - initial_capacity,
    capacity_growth_pct = (recent_capacity - initial_capacity) / initial_capacity * 100
  ) %>%
  arrange(desc(capacity_growth)) %>%
  drop_na()

# Calculate progress in electricity production
production_progress <- power_plant_data %>%
  select(country_long, matches("^generation_gwh_\\d+$")) %>%
  pivot_longer(
    cols = matches("^generation_gwh_\\d+$"),
    names_to = "year",
    values_to = "generation_gwh",
    values_drop_na = TRUE
  ) %>%
  mutate(year = as.numeric(gsub("generation_gwh_", "", year))) %>%
  group_by(country_long, year) %>%
  summarise(total_production = sum(generation_gwh, na.rm = TRUE), .groups = "drop") %>%
  arrange(country_long, year) %>%
  group_by(country_long) %>%
  summarise(
    initial_production = first(total_production),
    recent_production = last(total_production),
    production_growth = recent_production - initial_production,
    production_growth_pct = (recent_production - initial_production) / initial_production * 100
  ) %>%
  arrange(desc(production_growth)) %>%
  drop_na()

# Identify top-performing countries in capacity and production growth
top_capacity_countries <- capacity_progress %>%
  slice_max(capacity_growth, n = 10)

top_production_countries <- production_progress %>%
  slice_max(production_growth, n = 10)

# Plot capacity progress
p1 <- ggplot(top_capacity_countries, aes(x = reorder(country_long, capacity_growth), y = capacity_growth, fill = capacity_growth)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_viridis_c(option = "C") +
  labs(
    title = "Top 10 Countries by Installed Capacity Growth",
    x = "Country",
    y = "Capacity Growth (MW)",
    fill = "Growth (MW)"
  ) +
  theme_minimal()

if (is_html_output()) {
  ggplotly(p1)
} else {
  p1
}

# Plot production progress
p2 <- ggplot(top_production_countries, aes(x = reorder(country_long, production_growth), y = production_growth, fill = production_growth)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_viridis_c(option = "C") +
  labs(
    title = "Top 10 Countries by Electricity Production Growth",
    x = "Country",
    y = "Production Growth (GWh)",
    fill = "Growth (GWh)"
  ) +
  theme_minimal()

if (is_html_output()) {
  ggplotly(p2)
} else {
  p2
}
```


I trust more the code bellow:


```{r}
# 2. Identification of Countries Making the Most Progress
# Calculate progress in installed capacity and electricity production
# Reshape and calculate progress
progress <- power_plant_data %>%
  mutate(across(matches("^generation_gwh_\\d+$"), as.numeric)) %>%
  pivot_longer(
    cols = matches("^generation_gwh_\\d+$"),
    names_to = "year",
    values_to = "generation_gwh",
    values_drop_na = TRUE
  ) %>%
  group_by(country) %>%
  summarise(
    total_capacity = sum(capacity_mw, na.rm = TRUE),
    total_production = sum(generation_gwh, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    capacity_rank = rank(-total_capacity),
    production_rank = rank(-total_production)
  ) %>%
  arrange(capacity_rank, production_rank)


# View the progress summary
print(progress)

# Visualization: Progress by Top 10 Countries (Capacity and Production)
top_countries_progress <- progress %>%
  slice_head(n = 10)

ggplot(top_countries_progress, aes(x = reorder(country, -total_capacity), y = total_capacity, fill = country)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = "Top 10 Countries: Installed Capacity Progress",
    x = "Country",
    y = "Installed Capacity (MW)",
    fill = "Country"
  ) +
  coord_flip() +
  theme_minimal()

ggplot(top_countries_progress, aes(x = reorder(country, -total_production), y = total_production, fill = country)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = "Top 10 Countries: Electricity Production Progress",
    x = "Country",
    y = "Electricity Production (GWh)",
    fill = "Country"
  ) +
  coord_flip() +
  theme_minimal()
```

## Estimation of Future Capacities 

```{r}
# Prepare data for regression
capacity_trend <- power_plant_data %>%
  filter(!is.na(commissioning_year) & !is.na(capacity_mw)) %>%
  group_by(commissioning_year) %>%
  summarise(total_capacity = sum(capacity_mw, na.rm = TRUE), .groups = 'drop')

# Fit a linear model
capacity_model <- lm(total_capacity ~ commissioning_year, data = capacity_trend)

# Predict future capacities for the next 10 years
future_years <- data.frame(commissioning_year = seq(
  max(capacity_trend$commissioning_year) + 1, 
  max(capacity_trend$commissioning_year) + 10
))
future_years$total_capacity <- predict(capacity_model, newdata = future_years)

# Combine historical and predicted data
future_capacity_trend <- bind_rows(
  capacity_trend,
  future_years
)

# Visualization: Historical and Predicted Installed Capacity
p <- ggplot(future_capacity_trend, aes(x = commissioning_year, y = total_capacity)) +
  geom_line(color = "blue", size = 1) +
  geom_point(
    data = future_years, 
    aes(x = commissioning_year, y = total_capacity), 
    color = "red", size = 2
  ) +
  labs(
    title = "Historical and Predicted Installed Capacity",
    x = "Year",
    y = "Total Installed Capacity (MW)"
  ) +
  theme_minimal()

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```


## Temporal Analysis of Energy Production Trends (a revoir)

```{r}
# 1. Prepare Data for Time-Series Analysis

# Aggregate yearly production data for all power plants
yearly_production <- power_plant_data %>%
  select(matches("^generation_gwh_\\d+$")) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  pivot_longer(cols = everything(), names_to = "year", values_to = "total_production") %>%
  mutate(year = as.numeric(gsub("generation_gwh_", "", year))) %>%
  arrange(year)

# Convert to time-series object
ts_production <- ts(yearly_production$total_production, start = min(yearly_production$year), frequency = 1)

# View time-series data
print(yearly_production)

# Visualization: Yearly Production Trend
ggplot(yearly_production, aes(x = year, y = total_production)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "red") +
  labs(
    title = "Yearly Energy Production Trend",
    x = "Year",
    y = "Total Production (GWh)"
  ) +
  theme_minimal()
```

```{r}
# 2. Decompose Time-Series Data

# Decompose the time-series into components
decomposition <- decompose(ts_production, type = "additive")

# Plot decomposition
autoplot(decomposition) +
  labs(
    title = "Decomposition of Energy Production Time-Series",
    x = "Year",
    y = "Production (GWh)"
  )
```

```{r}
# 3. Forecast Future Production Using ARIMA

# Fit an ARIMA model
arima_model <- auto.arima(ts_production)

# Forecast for the next 10 years
forecast_production <- forecast(arima_model, h = 10)

# Plot forecast
autoplot(forecast_production) +
  labs(
    title = "Forecast of Future Energy Production",
    x = "Year",
    y = "Total Production (GWh)"
  )
```


```{r}
# 4. Detect Patterns: Seasonal and Cyclical Trends

# Highlight cyclical or seasonal patterns
cyclical_pattern <- ggplot(yearly_production, aes(x = year, y = total_production)) +
  geom_smooth(method = "loess", color = "blue", fill = "lightblue", alpha = 0.3) +
  labs(
    title = "Cyclical Trends in Energy Production",
    x = "Year",
    y = "Total Production (GWh)"
  ) +
  theme_minimal()

print(cyclical_pattern)

```
