## Mean Capacity of Power Plants by Primary Fuel Type

```{r}
# Mean, median, and maximum capacity of plants by fuel type (primary_fuel)
capacity_by_fuel <- power_plant_data %>%
  group_by(primary_fuel) %>%
  summarise(
    mean_capacity = mean(capacity_mw, na.rm = TRUE),
    median_capacity = median(capacity_mw, na.rm = TRUE),
    max_capacity = max(capacity_mw, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_capacity))

# View summary statistics
print(capacity_by_fuel)

# Mean, median, max capacity by fuel type
p <- ggplot(capacity_by_fuel, aes(x = reorder(primary_fuel, -mean_capacity), y = mean_capacity, fill = primary_fuel)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  labs(
    title = "Mean Capacity of Power Plants by Primary Fuel Type",
    x = "Fuel Type",
    y = "Mean Capacity (MW)"
  )

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```


## Comparison of Actual and Estimated Production

```{r}
# Gather data for both actual and estimated generation
actual_estimated_comparison <- power_plant_data %>%
  select(starts_with("generation_gwh_"), starts_with("estimated_generation_gwh_")) %>%
  summarise(across(everything(), ~ sum(.x, na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "year", values_to = "generation") %>%
  mutate(
    type = ifelse(grepl("estimated", year), "Estimated", "Actual"),
    year = gsub("generation_gwh_|estimated_generation_gwh_", "", year)
  )

print(actual_estimated_comparison)

p <- ggplot(actual_estimated_comparison, aes(x = year, y = generation, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = "Comparison of Actual and Estimated Production",
    x = "Year",
    y = "Total Generation (GWh)",
    fill = "Type"
  ) +
  theme_minimal()

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```


## Installed Capacity by Primary Fuel Type at Global Level

```{r}
# Average Installed Capacity by Primary Fuel Type at Global Level
avg_capacity_by_fuel <- power_plant_data %>%
  group_by(primary_fuel) %>%
  summarise(
    avg_capacity = mean(capacity_mw, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(avg_capacity))

# View summary table
print(avg_capacity_by_fuel)

# Visualization: Average Installed Capacity by Fuel Type
p <- ggplot(avg_capacity_by_fuel, aes(x = reorder(primary_fuel, -avg_capacity), y = avg_capacity, fill = primary_fuel)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  labs(
    title = "Average Installed Capacity by Primary Fuel Type",
    x = "Fuel Type",
    y = "Average Capacity (MW)",
    fill = "Fuel Type"
  ) +
  theme_minimal()

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```


# Average Production (GWh) by Primary Fuel Type or Country

```{r}
# Calculate average production by fuel type
average_production_by_fuel <- power_plant_data %>%
  mutate(
    total_production = rowSums(select(., matches("^generation_gwh_\\d+$")), na.rm = TRUE)
  ) %>%
  group_by(primary_fuel) %>%
  summarise(
    avg_production = mean(total_production, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(avg_production))

print(average_production_by_fuel)

# Visualization: Average Production by Fuel Type
p <- ggplot(average_production_by_fuel, aes(x = reorder(primary_fuel, -avg_production), y = avg_production, fill = primary_fuel)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  labs(
    title = "Average Production by Fuel Type (GWh)",
    x = "Fuel Type",
    y = "Average Production (GWh)",
    fill = "Fuel Type"
  ) +
  theme_minimal()

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```

## Average production by country

```{r}
# Calculate average production by country
average_production_by_country <- power_plant_data %>%
  mutate(
    total_production = rowSums(select(., matches("^generation_gwh_\\d+$")), na.rm = TRUE)
  ) %>%
  group_by(country) %>%
  summarise(
    avg_production = mean(total_production, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(avg_production))

print(average_production_by_country)

# Visualization: Average Production by Country (Top 10 Countries)
top_countries <- average_production_by_country %>%
  #top_n(10, wt = avg_production)
  head(10)

p <- ggplot(top_countries, aes(x = reorder(country, -avg_production), y = avg_production, fill = country)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  labs(
    title = "Average Production by Country (Top 10, GWh)",
    x = "Country",
    y = "Average Production (GWh)",
    fill = "Country"
  ) +
  theme_minimal()

if (is_html_output()) {
  ggplotly(p)
} else {
  p
}
```

## Clustering of Power Plants According to Performance and Characteristics

```{r}
# Prepare data for clustering
clustering_data <- power_plant_data %>%
  mutate(
    total_production = rowSums(select(., matches("^generation_gwh_\\d+$")), na.rm = TRUE),
    performance_ratio = total_production / (capacity_mw * 8760 / 1000) # Production as % of max capacity
  ) %>%
  select(primary_fuel, capacity_mw, total_production, performance_ratio) %>%
  filter(!is.na(capacity_mw) & !is.na(total_production) & !is.na(performance_ratio)) %>%
  mutate(primary_fuel = as.factor(primary_fuel)) # Convert primary_fuel to a factor

# Normalize numeric data
numeric_data <- clustering_data %>%
  select(capacity_mw, total_production, performance_ratio) %>%
  scale()

# Perform clustering using k-means
set.seed(123) # Ensure reproducibility
kmeans_result <- kmeans(numeric_data, centers = 3, nstart = 25)

# Add cluster labels to data
clustering_data <- clustering_data %>%
  mutate(cluster = as.factor(kmeans_result$cluster))

# Visualization: Clustering Results
ggplot(clustering_data, aes(x = capacity_mw, y = total_production, color = cluster, shape = primary_fuel)) +
  geom_point(size = 3) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    title = "Clustering of Power Plants Based on Performance and Characteristics",
    x = "Installed Capacity (MW)",
    y = "Total Production (GWh)",
    color = "Cluster",
    shape = "Primary Fuel"
  ) +
  theme_minimal()

```


## Impact of Age on Performance

To analyze the impact of age on the performance of power plants, we can calculate the age of each plant, evaluate its performance based on electricity generation (`generation_gwh_*` columns), and then visualize the relationship between age and performance.

Here’s how to approach this:

1. **Calculate Age of Power Plants:**
   - Age is the difference between the current year and the `commissioning_year`.

2. **Aggregate Performance:**
   - Sum up electricity generation across all available years for each plant.

3. **Filter Data:**
   - Exclude plants with missing or invalid `commissioning_year` or `generation_gwh_*` data.

4. **Visualize Relationship:**
   - Create scatter plots or line charts to show trends between plant age and performance.
   - Use `ggplot2` for visualization.


```{r}
# Set the current year
current_year <- as.integer(format(Sys.Date(), "%Y"))

# Calculate age and performance
plant_age_performance <- power_plant_data %>%
  filter(!is.na(commissioning_year)) %>% # Exclude rows with missing commissioning_year
  mutate(age = current_year - as.integer(round(commissioning_year))) %>% # Calculate age
  rowwise() %>%
  mutate(
    total_generation = sum(
      c_across(starts_with("generation_gwh_")), na.rm = TRUE
    )
  ) %>%
  ungroup() %>%
  filter(!is.na(total_generation), total_generation > 0, age >= 0) # Exclude invalid values

# Visualization: Scatter plot of Age vs. Performance
ggplot(plant_age_performance, aes(x = age, y = total_generation)) +
  geom_point(alpha = 0.6, color = "darkblue") +
  geom_smooth(method = "loess", color = "red", se = TRUE) +
  scale_y_log10(labels = scales::comma) + # Log scale for better visualization
  labs(
    title = "Impact of Plant Age on Performance",
    x = "Plant Age (Years)",
    y = "Total Electricity Generation (GWh)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none"
  )

# Additional Analysis: Group by Age and Summarize
age_summary <- plant_age_performance %>%
  group_by(age) %>%
  summarise(
    avg_generation = mean(total_generation, na.rm = TRUE),
    total_generation = sum(total_generation, na.rm = TRUE),
    plant_count = n()
  ) %>%
  arrange(age)

# Visualization: Average Generation by Age
ggplot(age_summary, aes(x = age, y = avg_generation)) +
  geom_line(color = viridis(1)) +
  labs(
    title = "Average Performance of Plants by Age",
    x = "Plant Age (Years)",
    y = "Average Electricity Generation (GWh)"
  ) +
  theme_minimal()
```

## Identify Older Plants with Low Performance

```{r}
# Define older plants (e.g., age > 100 years)
older_plants <- plant_age_performance %>%
  filter(age > 100)

# Define low performance as below the 25th percentile of total generation
low_performance_threshold <- quantile(older_plants$total_generation, 0.25, na.rm = TRUE)

# Identify older plants with low performance
older_low_performance <- older_plants %>%
  filter(total_generation < low_performance_threshold) %>%
  arrange(age, total_generation) # Sort for easier review

# View the results
print(older_low_performance)
```

```{r}
# Plot older plants with low performance
ggplot(older_low_performance, aes(x = age, y = total_generation)) +
  geom_point(color = "red", size = 3, alpha = 0.7) +
  labs(
    title = "Older Power Plants with Low Performance",
    subtitle = "Plants older than 100 years with generation below the 25th percentile",
    x = "Age of Plant (Years)",
    y = "Total Electricity Generation (GWh)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12, face = "italic")
  )
```







